Beliefs about the development of mental life

### 2021-10-16 paper/paper.Rmd
more anonymizing

---
title: "Beliefs about the development of mental life"
author: "[authors removed for anonymous review]"
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    fig_caption: yes
    toc: false
    number_sections: false
  #   reference_docx: word-styles-ref.docx
  # word_document:
bibliography: "../supplement/references.bib"
csl: "../supplement/apa.csl"
# cls: "../supplement/article.cls"
figsintext: yes
header-includes: 
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{float}
---

@@

Caregiving relationships with infants and children are among the most common, important, and complex human social interactions. Caring for a young child over time requires an adult to continually update their own mental representations of the child’s capacities as the child develops. In a series of large-scale studies, we assessed how US adults conceptualize the development of the human mind over the first five years of life. Exploratory factor analysis identified four categories of mental capacities that anchored participants’ representations of the developing human mind: _bodily sensation_ (e.g., hunger, pain), _negative affect_ (e.g., distress, frustration), _social connection_ (e.g., love, learning from others), and _cognition and control_ (e.g., planning, self-control). Participants perceived that these four aspects of mental life were present to different degrees at birth, followed different developmental trajectories, and were driven by different developmental mechanisms (such as biological "preprogramming," physical maturation, passive observation, and social learning). Adults’ perceptions of children’s mental capacities have important consequences for the well-being of children in their care—particularly in the first few years of life, when children’s communication skills are limited and caregivers must infer children’s thoughts, feelings, and needs. The current studies shed light on this fascinating and understudied aspect of "mind perception," in turn laying the foundation for theory-based interventions to encourage positive and responsive parenting behaviors.

>>>>

---
title: "Beliefs about the development of mental life"
author: '[authors removed for anonymous review]'
output:
  pdf_document:  default
  # word_document:
  #   reference_docx: word-styles-ref.docx
bibliography: ../supplement/references.bib
csl: ../supplement/apa.csl
figsintext: yes
header-includes:
- \usepackage{lineno}
- \linenumbers
- \usepackage{float}
---

@@

In a series of studies, we assessed how US adults conceptualize the development of the human mind over the first five years of life. Exploratory factor analysis identified four core capacities that anchored participants’ representations of the developing human mind: _bodily sensation_ (e.g., hunger, pain), _negative affect_ (e.g., distress, frustration), _social connection_ (e.g., love, learning from others), and _cognition and control_ (e.g., planning, self-control). Participants believed that these capacities were present to different degrees at birth, followed different developmental trajectories, and were driven by different developmental mechanisms (such as biological “preprogramming,” physical maturation, passive observation, and social learning). Caregiving relationships with infants and children are among the most common and most complex human social interactions, and adults’ perceptions of children’s mental capacities have important consequences for the well-being of children in their care—particularly in the first few years of life, when children’s communication skills are limited and caregivers must infer children’s rapidly developing thoughts, feelings, and needs. The current studies shed light on this fascinating and understudied aspect of “mind perception,” in turn laying the foundation for theory-based interventions to encourage positive and responsive parenting behaviors.

### 2021-10-16 supplement/supplement-main.Rmd
more anonymizing

\newpage
In a series of large-scale studies we assessed how US adults conceptualize the development of the human mind over the first five years of life. We identified four distinct suites of mental capacities that are perceived by US adults to develop over human infancy and early childhood, _bodily sensation_, _negative affect_, _social connection_, and _cognition and control_ (Study 1); charted how these different aspects of mental life are perceived to change over the first five years of a child's life (Studies 2-3); and explored the intuitive theories that underlie the observed differences in the perceived developmental trajectories of these four aspects of mental life (Study 3).

@@

Note: This is a fully reproducible manuscript, written in RMarkdown using the "knitr" package for R [@xie2021]. The full, editable script for all analyses, figures, and tables (here and in the main text) is available in the .Rmd file that generated this document, which can be found in the following GitHub repository: https://github.com/kgweisman/baby_mental_life_ms.

@@

```{r}
# NOTE: this is a bit of a circular process -- results files are generated in this .Rmd file, but they also need to be loaded in order to write the Extended Methods section. For this reason, results RDS files are included in the github repo.
s1_results <- readRDS("../outputs/s1_results.rds")
s2_results <- readRDS("../outputs/s2_results.rds")
s3_results <- readRDS("../outputs/s3_results.rds")

>>>>

\newpage
In a series of large-scale studies we assessed how US adults conceptualize the development of the human mind over the first five years of life. We identified four core capacities that are perceived by US adults to develop over human infancy and early childhood, _bodily sensation_, _negative affect_, _social connection_, and _cognition and control_ (Study 1); charted how these different aspects of mental life are perceived to change over the first five years of a child's life (Studies 2-3); and explored the intuitive theories that underlie the observed differences in the perceived developmental trajectories of these four aspects of mental life (Study 3).

@@

Note: This is a fully reproducible manuscript, written in RMarkdown using the "knitr" package for R [@xie2021]. The full, editable script for all analyses, figures, and tables (here and in the main text) is available in the .Rmd file that generated this document, which can be found in the following GitHub repository (anonymized for peer review): https://anonymous.4open.science/r/baby_mental_life_ms-C261.

@@

```{r}
# NOTE: this is a bit of a circular process -- results files are generated in this .Rmd file, but they also need to be loaded in order to write the Extended Methods section. For this reason, results RDS files are included in the github repo.
# To re-make results from scratch, comment out these lines, run this whole .Rmd file, and then uncomment these lines before knitting the supplement.

s1_results <- readRDS("../outputs/s1_results.rds")
s2_results <- readRDS("../outputs/s2_results.rds")
s3_results <- readRDS("../outputs/s3_results.rds")
